---
title: "Data preparation for modelling"
order: 1
---

## Data structure

The project aims to allow us to predict the value of a performance metric in future years (1, 2, 5 years, say) for each of the ICSs. Performance metrics could be impacted by demand and capacity metrics for the same years, or potentially years building up to that year.

To enable testing different models, the data structure should consider the following for each ICS and each year:

* a value for each target variable (performance metrics) - required
* a value for each predictor variable (demand and capacity metrics) - optional
* a lagged value (1 to x years lag) for target and predictor variable - optional
* inclusion of ICB code and year in for prediction variables?

## Splitting dataset

Data need to be split before training model. Some considerations are:

* Should the data be ordered prior to splitting (eg, should models be trained on earlier years, and tested on later years) - on occasion there won't be very long time series
* Train, validation and test set? Is there enough data to do this?

## Missing data

* How to accommodate independent providers in the metrics
* The input data has missing values. The data needs more investigation as to why it is missing (it could be really missing, eg, data not submitted, or it could be a zero, eg, a zero numerator for that metric). 

  - Some timeseries data for Trusts (eg beds available) stop being reported and it is difficult to understand why (eg RT5 stops being reported in Q1 2015 even though the trust still exists)
  - Some "Independent Sector Healthcare providers" stop reporting Sickness data mid-timeseries

## Other questions

* What to do with IMD? It is published by LSOA for a single year. Should the information for each ICS be applied for each year? How else to integrate deprivation into the analysis?

## Next steps

* check whether missing data are missing or zero (values could be zero but come through as missing because the numerator is zero)





## Results

This section documents the results of modelling steps for reference and the project progresses

### Linear model

* No lagged data
* selection pattern: `matches("^ESR|^Workforce|^Bed|age band|Year 6|GPPS")`
* data are ordered by time before splitting into train (2013-2018), val (2019) and test (2020)
* correlation threshold: 0.9
* impute missing with knn for all predictors

```{r}
# A tibble: 3 × 4
#   .metric .estimator  train validation
#   <chr>   <chr>       <dbl>      <dbl>
# 1 rmse    standard   0.0317     0.0432
# 2 rsq     standard   0.830      0.283 
# 3 mae     standard   0.0257     0.0353
```

### Linear model current and lag

* Current year and lagged data (including lagged target variable)
* selection pattern: `matches("^ESR|^Workforce|^Bed|age band|Year 6|GPPS")`
* data are ordered by time before splitting into train (2013-2018), val (2019) and test (2020)
* correlation threshold: 0.9
* impute missing with knn for all predictors

```{r}
# A tibble: 3 × 4
#   .metric .estimator  train validation
#   <chr>   <chr>       <dbl>      <dbl>
# 1 rmse    standard   0.0263     0.0363
# 2 rsq     standard   0.883      0.452 
# 3 mae     standard   0.0215     0.0281
```

### Linear model using only lagged data

* Only lagged data (including lagged target variable)
* selection pattern: `matches("^ESR|^Workforce|^Bed|age band|Year 6|GPPS")`
* data are ordered by time before splitting into train (2014-2018), val (2019) and test (2020)
* correlation threshold: 0.9
* impute missing with knn for all predictors

```{r}
# A tibble: 3 × 4
#   .metric .estimator  train validation
#   <chr>   <chr>       <dbl>      <dbl>
# 1 rmse    standard   0.0238     0.0368
# 2 rsq     standard   0.883      0.498 
# 3 mae     standard   0.0191     0.0254
```

### Linear model using only lagged data (without lagged target variable)

* Only lagged data (not including lagged target variable)
* selection pattern: `matches("^ESR|^Workforce|^Bed|age band|Year 6|GPPS")`
* data are ordered by time before splitting into train (2014-2018), val (2019) and test (2020)
* correlation threshold: 0.9
* impute missing with knn for all predictors

```{r}
# A tibble: 3 × 4
#   .metric .estimator  train validation
#   <chr>   <chr>       <dbl>      <dbl>
# 1 rmse    standard   0.0322     0.0449
# 2 rsq     standard   0.787      0.236 
# 3 mae     standard   0.0258     0.0368
```




### Random forest current and lag

* Current year and lagged data (including lagged target variable)
* selection pattern: `matches("^ESR|^Workforce|^Bed|age band|Year 6|GPPS")`
* data are ordered by time before splitting into train (2013-2018), val (2019) and test (2020)
* impute missing with knn for all predictors

```{r}
# A tibble: 3 × 5
#   .metric .estimator  train validation   test
#   <chr>   <chr>       <dbl>      <dbl>  <dbl>
# 1 rmse    standard   0.0166     0.0574 0.132 
# 2 rsq     standard   0.976      0.258  0.0839
# 3 mae     standard   0.0134     0.0493 0.125 
```